---
title: "Project dibimbing"
author: "dandi dirgantara"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Library
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
```
# Data Description
Feature Name	                                                Description
person_age	                                                  Age
person_income	                                                Annual Income
personhomeownership	                                          Home ownership
personemp                                                     length	Employment length (in years)
loan_intent	                                                  Loan intent
loan_grade	                                                  Loan grade
loan_amnt	                                                    Loan amount
loanintrate	                                                  Interest rate
loan_status	                                                  Loan status (0 is non default 1 is default)
loanpercentincome	                                            Percent income
cbpersondefaultonfile	                                        Historical default
cbpresoncredhistlength	                                      Credit history length


#Load Data
```{r}
df <- read.csv("G:\\My Drive\\dibimbing\\credit_risk_dataset.csv")
df
```

#Initial EDA
```{r}
str(df)
```

```{r}
summary(df)
```
- N/A value in  loan_int_rate, person_emp_length
- max age = 144 (weird)
- max emp_length = 123 (weird)

```{r}
df %>% 
  group_by(loan_status) %>% 
  summarise(count=n())
```
```{r}
#default rate is quite high 
(7108/32581) * 100
```


#Data Preparation
```{r}
df %>% is.na() %>% colSums()
```

```{r}
df %>% 
  filter(is.na(person_emp_length))
```

```{r}
df %>% 
  filter(is.na(loan_int_rate))
```

- we can drop na value from column int_rate (since it's not make sense there's people who apply loan with no loan interest) 
- we can drop na value from emp_length (since there's income data but no emp_length which not make sense)

```{r fig.width=10, fig.height=5}
df %>% 
  ggplot(aes(x=person_age, y=person_emp_length))+
  geom_point()+
  labs(title = "Age vs Employee Length") +
  scale_x_continuous(breaks = seq(20, 150, 25), 
                     limits = c(20,150))
```
- there's outliers from emp_length (young people but the emp_length >= 100 year)
- also with the age column (old people but with low emp_length)

```{r}
df %>% 
  filter (person_emp_length == 123)
```

```{r fig.width=10, fig.height=5}
df %>% 
  ggplot(aes(x=person_emp_length, y= person_age))+
  geom_point()+
  labs(title = "Age vs Employee Length") +
  scale_y_continuous(breaks = seq(20, 150, 25),
                     limits = c(20,150))
```
```{r}
df %>% 
  filter (person_age >= 90)
```

- so based on checking above we can drop these condition below:
  1. employ_length NA and >= 100
  2. person_age NA and juga >= 90
  
```{r}
df_clean <- df %>% 
  filter(!is.na(person_emp_length)) %>%
  filter(!is.na(loan_int_rate)) %>%
  filter(person_age < 90) %>%
  filter(person_emp_length < 100) %>% 
  filter(person_income < 2039784)

summary(df_clean)
```
```{r}
df_clean <- df_clean %>% relocate (loan_status, .after = cb_person_cred_hist_length)
```

```{r}
df_clean %>% 
  filter(person_income >= 1000000)
```
```{r}
df_clean %>% 
  ggplot(aes(x=person_income)) +
  geom_boxplot() +
  scale_x_continuous(label = scales::label_number(suffix = ' M', scale = 1e-6)) 
```


#EDA
```{r fig.width=30}
# library(psych)
pairs.panels(df_clean, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
) 
```
- from pair plot above we can take a quick overview for each column 
- feature that have high correlation is person_age & cbpresoncredhistlength also loan_grade & loan_int_rate (so later we can keep only one feature for our model)
- on this anaylisis i will take a look on several column only
- the consideration is 
  1. column with strong correlation with loan_status
  2. column that might be impactful for credit scoring (historical data)
  3. personal info that represent the behavior for this client
```{r fig.width=10}
df_clean %>%
  ggplot(aes(x=person_age))+
  geom_bar() +
  labs(title = "Person Age Distribution") +
  scale_x_continuous(breaks = seq(20, 90, 5),
                     limits = c(20,90))
```
- from that plot we know mostly our client is a young people since we have high number in age range 20 -30

```{r fig.width= 10 ,fig.height= 7}
df_clean %>% 
  ggplot(aes(x = person_age , y = person_income)) +
  geom_point(aes(color = factor(loan_status))) +
  labs(title = "Age vs Income by loan status",
       color = "Loan Status") +
  scale_x_continuous(breaks = seq(20, 90, 5),
                     limits = c(20,90)) +
  scale_y_continuous(label = scales::label_number(suffix = ' M', scale = 1e-6)) 
```
- from this plot we know that mostly people with higher income are rarely hit default rate
- and small notes that people with higher income mostly in young people

```{r fig.width= 10 ,fig.height= 7}
df_clean %>% 
  ggplot(aes(x = person_emp_length , y = person_income)) +
  geom_point(aes(color = factor(loan_status))) +
  labs(title = "Employment vs Income by loan status",
       color = "Loan Status") +
  # scale_x_continuous(breaks = seq(20, 90, 5),
                     # limits = c(20,90)) +
  scale_y_continuous(label = scales::label_number(suffix = ' M', scale = 1e-6)) 
```
- income and employment length not having clear linear relation meaning people with long employment lenght doesn't mean the income will be high as well


```{r fig.width= 10}
df_clean %>% 
  ggplot(aes(x=person_income, y=factor(loan_intent))) +
  geom_boxplot() +
  labs(title = "Income vs Loan Intention") +
  scale_x_continuous(breaks = seq(0, 2500000, 250000),
                     limits = c(0,2500000),
                     label = scales::label_number(suffix = ' M', scale = 1e-6)) 
```
- people with higher income are more likely apply the loan for home improvement purpose

```{r fig.width= 10}
df_clean %>% 
  ggplot(aes(x=person_age, y=factor(loan_intent))) +
  geom_boxplot() +
  labs(title = "Age vs Loan Intention") +
  scale_x_continuous(breaks = seq(20, 90, 5),
                     limits = c(20,90))
```
- highest average age is on home improvement loan intention (meaning people with age range in 28 mostly apply loan for home improvement)


```{r fig.width=12}
df_clean %>% 
  ggplot(aes(x=person_age, y=person_income)) +
  geom_point(aes(color = factor(loan_intent))) +
  labs(title = "Age vs Income by loan intention",
       color = "Loan Intention") +
  scale_x_continuous(breaks = seq(20, 90, 5),
                     limits = c(20,90)) +
  scale_y_continuous(breaks = seq(0, 2500000, 250000),
                     limits = c(0,2000000),
                     label = scales::label_number(suffix = ' M', scale = 1e-6)) +
  scale_colour_brewer(palette = 'Set3')
```
- the distribution for age data is right skewed, showing this dataset is populated by young people
- also with the income is skewed as well, showing that the distribution is not well balanced

```{r fig.width=12}
df_clean %>% 
  ggplot(aes(x=person_age, y=person_income)) +
  geom_point(aes(color = factor(person_home_ownership))) +
  labs(title = "Age vs Income by home ownership",
       color = "Loan Intention") +
  scale_x_continuous(breaks = seq(20, 90, 5),
                     limits = c(20,90)) +
  scale_y_continuous(label = scales::label_number(suffix = ' M', scale = 1e-6)) +
  scale_colour_brewer(palette = 'Set3')
```
- people with the mortgage home ownership seems have a higher income

```{r fig.width=10}
df_clean %>% 
  ggplot(aes(x=loan_int_rate, y=loan_intent )) +
  geom_boxplot() +
  labs(title = "Loan Intention vs Interest Rate",
       x = "Interest Rate",
       y = "Loan Intention")
```
- from this plot we know that loan for home improvement is the most high compare to the others


```{r fig.width= 10}
df_clean %>% 
  mutate(coun = n()) %>% 
  ggplot(aes(fill=cb_person_default_on_file, y=coun, x=loan_intent)) + 
  geom_bar(position="fill", stat="identity") +
  labs(title = "Historical default vs Loan Intention")
```
- we know the highest default is on home improvement loan purpose based on this plot

```{r}
df_clean %>% 
  ggplot(aes(x = loan_int_rate, y = factor(loan_status)))+
  geom_boxplot() +
  labs(title = "Loan Status vs Interest Rate",
       x = "Interest Rate",
       y = "Loan Status")
```

```{r}
df_clean %>% 
  ggplot(aes(x = person_income, y = factor(loan_status)))+
  geom_boxplot() +
  labs(title = "Loan Status vs Income",
       x = "Income",
       y = "Loan Status") +
  scale_x_continuous(label = scales::label_number(suffix = ' M', scale = 1e-6))
```

```{r}
df_clean %>% 
  ggplot(aes(x = loan_amnt, y = factor(loan_status)))+
  geom_boxplot() +
  labs(title = "Loan Status vs Loan Amount",
       x = "Loan Amount",
       y = "Loan Status") +
  scale_x_continuous(label = scales::label_number(suffix = ' K', scale = 1e-3))
```

- from 3 tables above we found some insight correlation between these variables
- higher interest rate is most likely will get the default loan
- a lower income also is most likely will get default
- also the higher loan amount most likely will get default
- so based on that 3 insight we can take new insight that higher interest rates, a lower income and higher loan amount increase the possibility to have a default.